Predicting Probability of Participants' Vote of Conservative and Liberty in 2025 Canadian election 

### 2024-04-19 paper/paper(1).qmd
Add files via upload

Founded at the inception of Canada in 1867, the Liberal Party has long been a prominent player in the nation's political arena. Rooted in the principles of liberalism—a cornerstone of Canadian political ideology—the party champions a balanced fiscal approach, advocating for judicious tax reductions to foster economic expansion, boost investment, and invigorate consumer spending. At the heart of the Liberal Party's platform is a commitment to economic vitality and job creation, with a focus on propelling Canada's growth and ensuring ample employment opportunities. The party also prioritizes the well-being of its constituents, pledging to sustain and enhance public services including healthcare, pensions, and education. Committed to environmental stewardship, the Liberal Party envisions a greener Canada, aiming for reduced carbon emissions and heightened conservation efforts.
Our analysis, drawn from five different polls, suggests a marginal preference for the Liberal Party over the Conservative Party.
This trend is echoed in Russell's 2023 article, which indicates a public perception of the Liberal Party's superior handling of these key issues.
Thus, our working hypothesis posits that the Liberal Party holds a slender advantage in the race for the upcoming election[@Brussell2023two].

@@

```{r}
#| label: fig-liberal-votes-and-consrvative_votes
#| fig-cap: "predicted probabilities for the Conservative and Liberal parties"
#| echo: false
cons_probs <- predict(step_cons, newdata = cleaned_survey_data, type = "response")
lib_probs <- predict(step_lib, newdata = cleaned_survey_data, type = "response")

@@

# Appendix {.unnumbered}

# Additional data details

# Model details {#sec-model-details}

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide', fig.width=8, fig.height=6}
library(bayesplot)
library(rstanarm)

prior_covariates <- normal(location = 0, scale = 2.5, autoscale = FALSE)
prior_intercept <- normal(0, 2.5)

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide', fig.width=8, fig.height=6}

prior_covariates <- normal(location = 0, scale = 2.5, autoscale = FALSE)
prior_intercept <- normal(0, 2.5)

@@

@fig-liberal-model \newpage

@fig-conservative-posterior \newpage

## Diagnostics

>>>>

Founded at the inception of Canada in 1867, the Liberal Party has long been a prominent player in the nation's political arena. Rooted in the principles of liberalism—a cornerstone of Canadian political ideology—the party champions a balanced fiscal approach, advocating for judicious tax reductions to foster economic expansion, boost investment, and invigorate consumer spending. At the heart of the Liberal Party's platform is a commitment to economic vitality and job creation, with a focus on propelling Canada's growth and ensuring ample employment opportunities. The party also prioritizes the well-being of its constituents, pledging to sustain and enhance public services including healthcare, pensions, and education. Committed to environmental stewardship, the Liberal Party envisions a greener Canada, aiming for reduced carbon emissions and heightened conservation efforts.
Our analysis, drawn from five different polls, suggests a marginal preference for the Liberal Party over the Conservative Party.
This trend is echoed in Russell's 2023 article, which indicates a public perception of the Liberal Party's superior handling of these key issues.
Thus, our working hypothesis posits that the Liberal Party holds a slender advantage in the race for the upcoming election[@russell2023two].

@@

```{r echo=FALSE, message=FALSE, warning=FALSE,}
#| label: fig-liberal-votes-and-consrvative_votes
#| fig-cap: "predicted probabilities for the Conservative and Liberal parties"
#| echo: false
cons_probs <- predict(step_cons, newdata = cleaned_survey_data, type = "response")
lib_probs <- predict(step_lib, newdata = cleaned_survey_data, type = "response")

@@

# Appendix

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide', fig.width=8, fig.height=6}
#| label: fig-liberal-votes
#| fig-cap: "Posterior and Perior check for Liberal Party"
#| echo: false
prior_covariates <- normal(location = 0, scale = 2.5, autoscale = FALSE)
prior_intercept <- normal(0, 2.5)

@@

```
@fig-liberal-votes shows that there is no big difference between posterior and prior model of liberal party. When discussing prior and posterior in the context of Bayesian statistics, the prior distribution represents our initial beliefs about the parameters before observing the data, while the posterior distribution is an updated belief that takes the observed data into account. The closer the posterior is to the prior, the less influence the observed data has had on our beliefs. This could mean that the data was consistent with our expectations or that the data wasn't particularly informative with respect to that parameter.

If we were to infer the relationship for the Liberal Party, without direct data it would be speculative. However, if the model parameters included variables that could be linked to political tendencies or support for the Liberal Party (such as demographics or location), one might analyze the posterior distributions of these parameters to determine how the observed data aligns with the initial assumptions. If the posterior distributions are very similar to the prior distributions, it would suggest that the data confirmed the initial beliefs about these factors' relationship to support for the Liberal Party without significant changes. This indicates that the model's predictions for how certain factors influence support for the Liberal Party were quite accurate to begin with.

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide', fig.width=8, fig.height=6}
#| label: fig-conservative-votes
#| fig-cap: "Posterior and Perior check for Conservative Party"
#| echo: false
prior_covariates <- normal(location = 0, scale = 2.5, autoscale = FALSE)
prior_intercept <- normal(0, 2.5)

@@

@fig-conservative-votes compares the posterior and prior estimates of parameters for a Bayesian hierarchical model, analyzing factors that influence votes for the conservative party. The left side  shows the estimated values of parameters after the model has been updated with the data; these are the revised beliefs about the parameters' values. The colors and positions of the dots represent the mean or median of the posterior distributions, and the vertical lines may indicate the uncertainty or credible intervals. The right side suggesting it would normally show the initial beliefs or assumptions about the parameters before any data was observed; these priors are specified by the analyst based on prior knowledge or assumptions. The lack of information on the "Prior" side indicates that this visualization is focusing on the posterior outcomes. Essentially, the relationship between posterior and prior is that the posterior is the updated belief after considering the evidence (data), whereas the prior represents the beliefs before seeing the data. The posterior is informed by and is an adjustment from the prior, incorporating the information gained from the observed data.
